\sysfull

### 2018-08-05 paper/asplos19/background.tex
revisions to background (incomplete)

Energy harvesting devices operate using energy extracted from sources such as
radio frequency transmissions and light. These devices elide tethered
power and batteries, instead collecting energy into a capacitor, operating when
sufficient charge accumulates, and turning off for recharge upon depletion of the buffer.
There are several energy harvesting battery-less hardware platforms. For
instance, computational RFIDs---open-source TI MSP430-based~\cite{wolverine}
WISP~\cite{wisp5} (with its variants such as
WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or
NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones
such as~\cite{medusa_farsens_2017}. Other platforms include ambient backscatter
tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less
phone~\cite{talla_imwut_2017}.

\textbf{Hardware Assumptions.} \sys is designed for the demands of existing and
future energy-harvesting platforms based around general purpose, commodity
computing components~\cite{wisp,msp430datasheet}. We assume a device with a
memory system that has fast, byte-addressable volatile and non-volatile memory;
in particular, our target platform, WISP~\cite{wisp}, is equipped with a
mixture of SRAM and FRAM. Our implementation leverages hardware support for
fast, bulk-copying between memories via DMA~\cite{msp430datasheet}. \sys does not
require architectural additions to commodity processors as
in~\cite{su_date_2017,hicks_isca_2017,quickrecall,nvp}.

@@

\begin{figure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-example.pdf}
        \caption{WAR-affected code}
        \label{fig:war-example}
    \end{subfigure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-execution.pdf}
        \caption{Incorrect execution}
        \label{fig:war-execution}
    \end{subfigure}
    \caption{WAR dependency example. \texttt{NV} marks non-volatile variables,
    while \texttt{checkp()} is a checkpoint of volatile state.}
    \label{fig:war}
\end{figure}

@@

A power failure clears volatile state (e.g., registers and SRAM) while
non-volatile memory (e.g., FRAM) persists. Upon a power failure, control flows
to a prior point in the execution: by default, to the beginning of {\tt
main()}. Early intermittent systems preserved progress by periodically
checkpointing volatile execution context to non-volatile
memory~\cite{mementos}, sometimes using hardware
support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}. 
%
Checkpointing volatile state alone does not ensure data consistency when the
system can directly manipulate non-volatile memory~\cite{mspcdino}.
Precisely, data can get inconsistent when code includes a
\emph{write-after-read} (WAR) dependency between operations that manipulate
non-volatile memory.
% because it allows {\em writes} from a failed attempt to
% persist across failure and be read by the {\em reads}~\cite{ratchet,dino,alpaca}.
Fig.~\ref{fig:war} illustrates how state can become
inconsistent in an intermittent execution using a simple example of an average
operation over an array of integers.
The non-volatile variable \texttt{sum} introduces the WAR, being read and written sequentially
by the increment operation.
If a power failure occurs right before updating the non-volatile index \texttt{i} (\ref{fig:war-execution}, Line 7), \texttt{sum} gets erroneously incremented twice consecutively by the same array element (\ref{fig:war-execution}, Lines 6 and 9).

>>>>

Energy harvesting devices operate using energy extracted from ambient sources, e.g. radio frequency transmissions, light. These devices elide tethered power and batteries, instead collecting energy into a capacitor, operating when sufficient charge accumulates, and turning off for recharge upon depletion of the buffer. There are several energy harvesting battery-less hardware platforms. For instance, computational RFIDs---open-source TI MSP430-based~\cite{wolverine} WISP~\cite{wisp5} (with its variants such as WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones such as~\cite{medusa_farsens_2017}. Other platforms include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less phone~\cite{talla_imwut_2017}.

\textbf{What Hardware does \sys Target.} \sys is designed for the demands of existing and future energy-harvesting platforms based around general purpose, commodity computing components~\cite{wisp,msp430datasheet}. \sys targets a device with a
memory system that has fast, byte-addressable volatile and non-volatile memory; in particular, our target platform, WISP~\cite{wisp}, is equipped with a mixture of SRAM and FRAM. \sys leverages hardware support for fast, bulk-copying between memories via DMA~\cite{msp430datasheet}. \sys does not require architectural additions to commodity processors as in~\cite{su_date_2017,hicks_isca_2017,quickrecall,nvp}.

@@

\begin{figure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-example.pdf}
        \caption{WAR-affected code}
        \label{fig:war-example}
    \end{subfigure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-execution.pdf}
        \caption{Incorrect execution}
        \label{fig:war-execution}
    \end{subfigure}
    \caption{WAR dependency example. \texttt{NV} marks non-volatile variables, while \texttt{checkp()} is a checkpoint of volatile state. \task{replace checkp(); with checkpoint();}}
    \label{fig:war}
\end{figure}

@@

A power failure clears volatile state (e.g., registers and SRAM) while non-volatile memory (e.g., FRAM) persists. Upon a power failure, control flows to a prior point in the execution: by default, to the beginning of {\tt main()}. Early intermittent systems preserved progress by periodically checkpointing volatile execution context to non-volatile memory~\cite{mementos}, sometimes using hardware support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}. Checkpointing volatile state alone does not ensure data consistency when the system can directly manipulate non-volatile memory~\cite{mspcdino}. Precisely, data can get inconsistent when code includes a \emph{write-after-read} (WAR) dependency between operations that manipulate non-volatile memory. Fig.~\ref{fig:war} illustrates how state can become inconsistent in an intermittent execution using a simple example of an average operation over an array of integers.
The non-volatile variable \texttt{sum} introduces the WAR, being read and written sequentially by the increment operation. If a power failure occurs right before updating the non-volatile index \texttt{i} (\ref{fig:war-execution}, Line 7), \texttt{sum} gets erroneously incremented twice consecutively by the same array element (\ref{fig:war-execution}, Lines 6 and 9).
